---
title: "CoronaNet Research Project"
author: Joan Barceló, Cindy Cheng,  Allison Spencer Hartnett, Robert Kubinec,Luca Messerschmidt
date: "https://coronanet-project.org"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE,tinytex.verbose = TRUE)

# to compile with pandoc
# do brew install pandoc pandoc-citeproc
# then add this line to a .Renviron file in the project folder:
# PATH="/usr/local/Cellar/pandoc/2.9.2.1/bin/:${PATH}"

require(dplyr)
require(tidyr)
require(ggplot2)
require(lubridate)
require(stringr)
require(kableExtra)
require(ggrepel)
library(readr)
library(knitr)
library(readxl)
require(gghighlight)


library(devtools)
library(igraph)
library(remotes)
#remotes::install_github("wjrl/RBioFabric")
library(RBioFabric)
library(xtable)
library(Hmisc)
# let's load some data!
# run RCode/cleanData/cleanQualtrics_short.R

clean_data <- readRDS("../data/CoronaNet/coranaNetData_clean.rds") %>% 
  mutate(date_announced=mdy(date_announced)) %>% 
  filter(date_announced<(today()-days(5)),!is.na(init_country),is.na(init_other),is.na(target_other) | target_other=="")

# run /RCode/slack/slackAnalytics.R"

slack <- readRDS("../data/slack/corona_govt_response_slack_latest_clean.rds")

# run /RCode/country_regions.R
regions_df = read.csv( "../data/regions/country_regional_groups_concordance.csv", stringsAsFactors = FALSE)
countries = read.csv('../data/regions/all_countries.csv', stringsAsFactors = FALSE)
```

# CoronaNet COVID-19 Government Response Database Overview 

Database of government policy actions directed toward addressing the COVID-19 pandemic

- Coverage
    + 193 countries
    + 6,816 policy events
    + Time range: December 31, 2019 to April 14, 2020
- Availability
  + Publicly available at https://coronanet-project.org
  + Updated daily with a 5-day lag for record validity checking
- Collaborators
  + Project Investigators: Joan Barceló, Cindy Cheng,  Allison Spencer Hartnett, Robert Kubinec,Luca Messerschmidt
  + Over 200 Research Assistants across 18 time zones collecting data
  
#  Variables 


- The type of government policy implemented (e.g. quarantine [16 total] )
- The level of government initiating the action (e.g. national, provincial )
- The geographical target of the policy action, if applicable (e.g. national, provincial)
- The human or material target of the policy action, if applicable (e.g. travelersf)
- The directionality of the policy action, if applicable (e.g. inbound, outbound, both)
- The mechanism of travel that the policy action targets, if applicable (e.g. flights )
- The compliance with the policy action (e.g. mandatory, voluntary)
- The enforcer of the policy action (e.g. national government, military)
- The timing of the policy action (e.g. date announced, date implemented)


```{r overtime,fig.cap="Cumulative Incidence of Policy Event Types Over Time"}

clean_data %>% 
  filter(!is.na(type)) %>% 
  group_by(type,date_announced) %>% 
  summarize(Policies=length(unique(record_id))) %>% 
  arrange(type,date_announced) %>% 
  mutate(Policies=cumsum(Policies)) %>% 
  ungroup %>% 
        mutate( type=recode(type,
                     `Public Awareness Campaigns`="Public\nAwareness\nCampaigns",
                     `External Border Restrictions`="External\nBorder\nRestrictions",
                     `Other Policy Not Listed Above`="Other",
                     `Restriction of Non-Essential Businesses`="Restriction of\nNon-Essential\nBusinesses",
                     `Restrictions of Mass Gatherings`="Restrictions of\nMass Gatherings",
                     `Restriction of Non-Essential Government Services`="Restriction of\nNon-Essential\nGovernment Services",
                     `Declaration of Emergency`="Declaration of\nEmergency",
                     `Internal Border Restrictions`="Internal\nBorder Restrictions",
                     `External Border Restrictions`="External\nBorder Restrictions",
                     `Public Awareness Campaigns`="Public\nAwareness Campaigns",
                     `New Task Force or Bureau`="New Task Force")) %>% 
  ggplot(aes(y=Policies,x=date_announced)) +
  geom_area() +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        strip.background = element_blank()) +
  xlab("") +
  facet_wrap(~type)

```


