---
title: "POMEPS Contribution"
author: "Robert Kubinec"
date: "4/13/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = F,message = F)

require(dplyr)
require(tidyr)
require(ggplot2)
require(lubridate)
require(stringr)

# load data

clean_data <- readRDS("../data/CoronaNet/coranaNetData_clean.rds")%>% 
  mutate(date_announced=mdy(date_announced)) %>% 
  filter(date_announced<(today()-days(5)),!is.na(init_country),is.na(init_other),is.na(target_other) | target_other=="")

```

The CoronaNet government response database (principal investigators Joan Barceló, Cindy Cheng, Allison Spencer Hartnett, Robert Kubinec and Luca Messerschmidt) is an effort to collect detailed data on policies taken by countries around the world in response to the COVID-19 pandemic. This project harnesses the efforts of more than 150 research assistants in 18 time zones to track countries across the world. The data is designed to permit inter-country comparisons across a range of policy types, both domestic and international. We also track the targets of policies so that we can identify dyadic and other network-based relationships, such as who implements travel restrictions against whom. The data will be continuously updated during the course of the pandemic, and are available via our website (link: https://coronanet-project.org/) and Github page (https://github.com/saudiwin/corona_tscs).

The data can also shed some light on the actions taken by Middle Eastern and North African countries. The plot below shows the number of national restrictions by policy type for the 14 countries we have managed to collect data on thus far (Algeria, Bahrain, Egypt, Iraq, Iran, Morocco, Saudi Arabia, Oman, Sudan, Syria, Turkey, Tunisia, the United Arab Emirates and Yemen). While this plot only counts the total number of policies, not their severity, it is clear that countries in the region have been most active at external border restrictions, while social distancing policies and restrictions of mass gatherings and non-essential businesses arrived relatively late. 


```{r mepolicy}

me <- c("Algeria","Bahrain","Egypt","Iraq","Iran","Morocco","Saudi Arabia","Oman","Sudan","Syria","Turkey","Tunisia","United Arab Emirates","Yemen")

just_me <- filter(clean_data,init_country %in% me)

outplot <- just_me %>%   filter(!is.na(type)) %>% 
  group_by(type,date_announced) %>% 
  summarize(Policies=length(unique(record_id))) %>% 
  arrange(type,date_announced) %>% 
  mutate(Policies=cumsum(Policies)) %>% 
  ungroup %>% 
        mutate( type=recode(type,
                     `Public Awareness Campaigns`="Public\nAwareness\nCampaigns",
                     `External Border Restrictions`="External\nBorder\nRestrictions",
                     `Other Policy Not Listed Above`="Other",
                     `Restriction of Non-Essential Businesses`="Restriction of\nNon-Essential\nBusinesses",
                     `Restrictions of Mass Gatherings`="Restrictions of\nMass Gatherings",
                     `Restriction of Non-Essential Government Services`="Restriction of\nNon-Essential\nGovernment Services",
                     `Declaration of Emergency`="Declaration of\nEmergency",
                     `Quarantine/Lockdown`="Quarantine/\nLockdown",
                     `Internal Border Restrictions`="Internal\nBorder Restrictions",
                     `External Border Restrictions`="External\nBorder Restrictions",
                     `Public Awareness Campaigns`="Public\nAwareness Campaigns",
                     `New Task Force or Bureau`="New Task Force")) %>% 
  distinct %>% 
  ggplot(aes(y=Policies,x=date_announced)) +
  geom_area() +
  theme_minimal() +
  scale_x_date(date_breaks = "1 month",date_labels="%d %b") +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(size=7),
        axis.text.x=element_text(size=7)) +
  xlab("") +
  facet_wrap(~type)

ggsave("me_plot.png",scale=1.1)

```

```{r border}

border <- filter(just_me,type=="External Border Restrictions")
iran <- filter(border, grepl(x=event_description,pattern="Iran") | target_country=="Iran")
```

It would appear that at least part of the early external border restrictions are due to the early outbreak in Iran. There are `r length(unique(iran$record_id))` policies targeting Iran, the earliest of which was announced on `r min(iran$date_announced,na.rm=T)`. However, as of April 8th, the top countries targeted with specific border restrictions were primarily outside of the region, as this table shows:

```{r restrict}

border %>% 
  group_by(target_country) %>% 
  summarize(n=length(unique(record_id)),
            n_unique=length(unique(init_country))) %>% 
    arrange(desc(n)) %>% 
  select(`Targeted Country`="target_country",`Count of Policy Restrictions`=n,`Number of Countries with Restriction`=n_unique) %>% 
  knitr::kable(.)

```

As of table shows, by April 8th, 12 out of the 14 countries in our data had completely closed off borders to any external flights. By comparison, the top most targeted countries were Iran and Italy, followed by China, Bahrain, Kuwait and Iraq. This regional breakdown suggests that countries were very aware of the early threat posed by outbreaks in neighboring states, and were willing to accept the diplomatic consequenes of banning travelers within the region. However, these early patterns are now subsumed by the fact that most borders in the region are closing completely, as is happening everywhere in the world.
